---
title: "exo_ggplot2"
author: "Florèn HUGON"
date: "4 novembre 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Lire les donnees
```{r}
devtools::load_all()
library(tidyverse) #autorise l'utilisation des pipes (%>%)

#bien mettre le here::here dans le Rmd
pantheria <- readr::read_delim(here::here("data", "pantheria-traits", "PanTHERIA_1-0_WR05_Aug2008.txt"), delim = "\t")
pantheria
```


### Nettoyer le dataset
Transformation en facteur de l'ordre et de la famille
```{r}
dat <- pantheria %>%
  mutate(order = as_factor(MSW05_Order), family  = as_factor(MSW05_Family)) %>% #ajoute ces deux colonnes
  janitor::clean_names() #permet de nettoyer les noms
```

Value -999 devient des NA, renommer les colonnes
```{r}
dat <- pantheria %>%
  mutate(order = as_factor(MSW05_Order), family  = as_factor(MSW05_Family)) %>% #ajoute ces deux colonnes
  rename(adult_bodymass = "5-1_AdultBodyMass_g",
         dispersal_age  = "7-1_DispersalAge_d",
         gestation      = "9-1_GestationLen_d",
         homerange      = "22-2_HomeRange_Indiv_km2",
         litter_size    = "16-1_LittersPerYear",
         longevity      = "17-1_MaxLongevity_m") %>%
         select(order, family, longevity, homerange, adult_bodymass, litter_size, dispersal_age, gestation) %>%
  na_if(-999) #met un NA si valeur = -999
dat
```


### Exploration des donnees
Consignes : How many observations in family? In order? What is the mean home range by family? Standard deviation? The sample size?

```{r}
dat %>%
  filter(!is.na(homerange)) %>% #permet d'enlever toutes les lignes avec NA sur les homerange, c'est mieux
  group_by(family) %>%
  summarise(moy=mean(homerange), sd=sd(homerange), n=n()) #n donne le nbr d'obs dans le groupe homerange(+ efficace)

```


### Premier graphe
Consignes : Group by family, and select only those with more than 100 observations. Make a bar chart of the number of observations per family. Order the bars by increasing number of observations. Add a title and labels to the X and Y axes.

```{r}
dat_family <- dat %>%
  group_by(family) %>%   # group by family
  mutate(n = n()) %>%    # calculate number of entries per family
  filter(n > 100) %>%    # select only the families with more than 100 entries
  
  ggplot() + #je mets d'abord mes donnees
  aes(x = n, y = fct_reorder(family, n)) +  #representation du nbr d'obs par famille
  geom_col() + #geometrie bar chart
  #coord_flip() pour echanger x et y
  xlab("Family") + # add label for X axis
  ylab("Counts") + # add label for Y axis
  ggtitle("Number of entries per family") # add title
dat_family
```


### Deuxième graphe
Consignes : Filter out missing values in both litter size and longevity. Group by family, and select only those with more than 50 observations. Make a scatter plot of litter size as a function of longevity. Add a linear regression fit for each family. Split the plot in several scatter plots, one for each family. Add a title and labels to the X and Y axes.

```{r}
dat %>% filter(!is.na(litter_size), !is.na(longevity)) %>% #enlever les lignes avec littersize et longevity=NA
  group_by(family) %>% #groupe par famille
  mutate(n = n()) %>%  #calcul nbr d'entree par famille
  filter(n > 10)  %>%  #conserve que celle avec au moins 10obs

  ggplot()+
  aes(x=longevity, y=litter_size, col=family)+ #une couleur par famille
  geom_point()+
  geom_smooth(method="lm", se=F)+
  #regression lineaire par famille (car on mit dans aes, col=family), se=F permet d'enlever l'IC
  labs(x="longevity (month)", y="litter size")+
  facet_wrap( ~ family, nrow = 3)+             #permet de splitter le plot par famille
  guides(color=guide_legend(title = "Family")) #ca marche mais c'est lourd
```

